Untrusted B2B Collaboration using the Blockchain

### 2018-07-18 paper/paper.tex
Start with implementation section

\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

@@

@@

>>>>

\usepackage{graphicx}
\usepackage{listings}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\input{solidity-highlighting.tex}

@@

Given the architecture, we will now discuss the implementation of a) the contracts used in the system, and b) the tools for interacting with the system.

@@

For taking part in the process collaboration, each company needs to deploy a factory contract that contains the contract definition of the company's instance contract.
This closely follows the factory contract pattern used in \cite{weber2016untrusted}, and described as Pattern 14 in \cite{xu2018pattern}.

- Split between choreography intiator and other participants
- has following interface:
  - constructor, passing the address of the access contract (needs to deployed first, details in next section)
  - setters for dependent factories
  - createInstance()
  - getAccessAddress()
  - isInstance(id, address)
- has following internal impl
  - guard instance creation until all dependent factories are set
  -> hold list of addresses of dependent factories
  - hold address of access contract
  - hold mapping from uuid -> address, identifying its own instances for a given choreography id
  - if initiator: generate uuid on createInstance

The factory contract source code also contains the public interfaces of
  - the access contract (same for every participant)
  - remote factory contracts (if receiving messages from remote, for verification)
  - remote instance contracts (if sending messages to remote instance)
as well as the complete definition of the instance contract source code.

The instance contract provides the following interface
  - a constructor where the local access contract, the choreography id and all dependency addresses can be passed
  - a function for each sending action (task or event), called from process engine
  - a function for each receiving action (task or event), called from other instance contracts
and the following internal impl
  - hold the choreography id
  - hold the addresses of its own factory (for selfdestruction) and its own access contract (for local verification)
  - hold the addresses of remote factory and access contract (again, if receiving messages from remote)
  - hold the addresses of remote instance contract (again, if sending messages)
  - add guards to each sending action to verify the local sender
  - add guards to each receiving action to verify the remote sender
  - implement events and emit them to the blockchain on receive action, that can be watched from external triggers
  - call remote receive functions on local send action
  - selfdestruct on the last send or receive action of the instance

Guards = access checks (Pattern 13)

See appendix for example % TODO

\subsubsection{Access Contract}

Each company deploys an access contract that holds information about the roles and permissions of the company's employees.
Having permission data encapsuled in a separate access contract follows pattern 12, the Data Contract pattern, from \cite{xu2018pattern}.
It allows decoupling the roles and permissions from the actual contract instances, allowing for a simple modification at runtime.
Access contracts can be deployed per process, or even once per company as a single source of truth for all processes.

To allow the administration of user permissions, access contracts provide the following interface:
\begin{itemize}
  \item 
    \texttt{assignRoleToUser(userAddress, role)}
  \item
    \texttt{removeRoleFromUser(userAddress, role)}
  \item
    \texttt{assignTaskToRole(task, role)}
  \item
    \texttt{removeTaskFromRole(task, role)}
\end{itemize}
The access to these functions should be guarded, allowing only a specific administrator address to call them.

Verifying a users permission to execute a certain task can be achieved by calling \texttt{isAuthorized(userAddress, task)}.
\newline

Given this interface, there are multiple possible implementation approaches for access contracts.
As it was our main focus to develop a running prototype, and not to provide an optimized, production-ready system, we decided on a simple mapping-based implementation:
\begin{lstlisting}[
  caption=Mappings in our access contract implementation,
  language=Solidity,
  basicstyle=\small
]
pragma solidity ^0.4.23;

contract Access {
    mapping(string => string) taskToRole;
    mapping(string => uint) roleToIndex;
    mapping(address => uint) addressToAccessBitmask;
    ...
}
\end{lstlisting}
The first mapping stores the tasks that can be executed by a specific role (cf. pools and lanes, and their contained tasks, in BPMN diagrams).
The second mapping assigns each role an index in the access bitmask, i.e., if the bit at the given index is set, the user holds the corresponding role.
The third mapping assigns such an access bitmask to each user.

Storing large amounts data on the blockchain is expensive by design.
As the mappings grow in size with each user, role and task that is part of some process in the company, our implementation is probably not suited for real-world application.
Finding optimized, less costly (both in computation cost and storage cost) approaches is part of future work.
One other approach we thought of is based on Pattern 5 of \cite{xu2018pattern} - depicting real world permissions as tokens, and "paying" with these tokens on function call to verify one's role.

@@

\section{Appendix}
% TODO highlight different parts as described in section Implementation? or remove?
\begin{lstlisting}[
  caption=Example source code of the Bulk Buyer's factory contract,
  language=Solidity,
  basicstyle=\small,
  xleftmargin=-3cm,
  xrightmargin=-3cm]
pragma solidity ^0.4.24;

contract Manufacturer {
    function receiveOrder(address) public;
    function setBulkBuyer(address) public;
}

contract ManufacturerFactory {
    function createInstance(uint) public returns(address);
    function getAccessAddress() public returns(address);
    function isInstance(uint, address) public view returns(bool);
}

contract Access {
    function isAuthorized(address, string) public returns(bool);
}

contract BulkBuyer {
    event ProductReceived();
    event StartOfProductionReceived();
    event Selfdestructed();

    address _factory;
    Access _localAccess;
    uint _id;

    ManufacturerFactory _manufacturerFactory;
    Access _manufacturerAccess;
    Manufacturer _manufacturer;

    modifier manufacturerAuthorized(address sender, string taskName) {
        require(_manufacturerFactory.isInstance(_id, msg.sender));
        require(_manufacturerAccess.isAuthorized(sender, taskName));
        _;
    }

    modifier localAuthorized(string taskName) {
        require(_localAccess.isAuthorized(msg.sender, taskName));
        _;
    }

    constructor(
        address localAccess,
        uint counter,
        address manufacturerFactory, address manufacturerAccess, address manufacturerInstance
    ) public {
        _factory = msg.sender;
        _localAccess = Access(localAccess);
        _id = counter;

        _manufacturerFactory = ManufacturerFactory(manufacturerFactory);
        _manufacturerAccess = Access(manufacturerAccess);
        _manufacturer = Manufacturer(manufacturerInstance);
    }

    function placeOrder() public localAuthorized("receiveOrder") {
        _manufacturer.receiveOrder(msg.sender);
    }

    function receiveStartOfProduction(address sender) 
      public manufacturerAuthorized(sender, "receiveStartOfProduction")
    {
        emit StartOfProductionReceived();
    }

    function receiveProduct(address sender) 
      public manufacturerAuthorized(sender, "receiveProduct")
    {
        emit ProductReceived();
        emit Selfdestructed();
        selfdestruct(_factory);
    }
}

contract BulkBuyerFactory {
    event BulkBuyerInstanceCreated(address instanceAddress);

    address _accessAddress;
    uint _counter;

    address _manufacturerAddress = address(0);

    mapping(uint => address) _instances;

    modifier initialized {
        require(_manufacturerAddress != address(0));
        _;
    }

    constructor(address accessAddress) public {
        _accessAddress = accessAddress;
        _counter = 0;
    }

    function setManufacturer(address manufacturerAddress) public {
        _manufacturerAddress = manufacturerAddress;
    }

    function createInstance() public initialized returns(address) {
        _counter = _counter + 1;

        ManufacturerFactory manufacturerFactory = ManufacturerFactory(_manufacturerAddress);
        address manufacturerAccess = manufacturerFactory.getAccessAddress();
        address manufacturerInstance = manufacturerFactory.createInstance(_counter);

        BulkBuyer b = new BulkBuyer(
            _accessAddress,
            _counter,
            manufacturerFactory, manufacturerAccess, manufacturerInstance
        );
        _instances[_counter] = address(b);
        emit BulkBuyerInstanceCreated(b);
        return b;        
    }

    function getAccessAddress() public view returns(address) {
        return _accessAddress;
    }

    function isInstance(uint id, address instanceAddress) public view returns(bool) {
        return _instances[id] == instanceAddress;
    }
}
\end{lstlisting}
